---
title: "HIPOVITAMINOSIS D AND IRON DEFICIENCY IN CHILDREN, A NEW APPROACH TO ITS UNEXPLAINED RELATIONSHIP WITH OSTEOARTICULAR PAIN"
author: "María Amelia Muñoz Calonge, Marcos Marvá Ruiz, David Varillas Delgado, José Jiménez Martínez and Manuel Rodríguez Zapata"
date: '2023-05-22nd'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document gathers the code we used to the statistical analysis of the data presented in the above referred manuscript.




### Data reading
```{r}
rm(list = ls())

dfcase0 = read.table(file = "cases_ms.csv", header = T, dec = ".", sep = ";")

dfcontrol0 = read.table(file = "controls_ms.csv", header = T, dec = ".", sep = ";")
```


### Controls:
 
 * 2 with no analiticals
 * 1 with no age
 * one empty line
 * one with iron < 10, I have set 7


### Setting baseline values

We first calculate the baseline values for `iron`, `ferritin` and `vitamin D` for both the case and control group. Note that

 * Not all the variables have been  measured in all the individuals because of occasional failures in the laboratory. 
 * Not all the individuals in the case group attended the second call and made the second blood test.
 
Thus: 

 * We use all the available data to set the baseline value of each variable for each group (case and control). We perform a t test to get a 95% confidence interval for the mean value.
 * Once the baseline values are set, we carry in analysing those case individuals that attended twice to the doctor.



#### Case group baseline values

We set the baseline value of each variable by calculating the corresponding confidence interval with the case individuals

```{r}
t.test(dfcase0$vitaminD, mu = 30, alternative = "two.sided", na.action(dfcase0$vitaminD))

t.test(dfcase0$iron, mu = 30, alternative = "two.sided",na.action(dfcase0$iron))

t.test(dfcase0$ferritin, mu = 30, alternative = "two.sided",na.action(dfcase0$ferritin))
```




Note that the case group 95% confidence interval for the mean value of `vitamin D` is below the threshold value 30 (deficiency).


#### Control group baseline values

```{r}
t.test(dfcontrol0$vitaminD, mu = 30, alternative = "two.sided", na.action(dfcontrol0$vitaminD))

t.test(dfcontrol0$iron, mu = 30, alternative = "two.sided",na.action(dfcontrol0$iron))

t.test(dfcontrol0$ferritin, mu = 30, alternative = "two.sided",na.action(dfcontrol0$ferritin))
```




### Data cleaning

We carry on by removing from the case group those individuals that did not complete the follow up 
 
```{r}
cases_complete = which(!is.na(dfcase0$iron2) & !is.na(dfcase0$ferritin2) &  !is.na(dfcase0$vitaminD2))
cases = dfcase0[cases_complete , ]
```


There are also individuals in the case group with no blood test records 

```{r}
dfcontrol0[is.na(dfcontrol0$vitaminD) , ]
```

that we, obviously, remove

```{r}
controls = dfcontrol0[!is.na(dfcontrol0$vitaminD), ]
```



### Data exploration

We first plot scatter plots and box plots for each variable and patient group.

```{r, fig.align='center', fig.height=5, fig.width=9}
par(mfrow = c(1,3), mar = c(4, 3, 1, 0), cex.axis = 1.5, cex.main = 1.5)

## diagram, VITAMIN D in groups C0, C1 and C2
dfVit = data.frame(vitD = c(cases$vitaminD, cases$vitaminD2, controls$vitaminD), 
                   group = c(rep("C1", length(cases$vitaminD)), 
                             rep("C2", length(cases$vitaminD2)), 
                             rep("C0", length(controls$vitaminD)))
) 
boxplot(dfVit$vitD ~ dfVit$group, horizontal = T, main = "Vitamin D", 
        xlab = "ng/ml", ylab = "")
stripchart(dfVit$vitD ~ dfVit$group, pch = c(15, 16, 17), add = T,
           col = gray.colors(3, start = 0.1, end = 0.55, gamma = 3.2, 
                             rev = TRUE), cex = 1.5,
           method = "jitter", xlim=c(0,max(dfVit$vitD)+1), 
           main = "Vitamin D")

# lines indicating vitamin D severe deficiency and vitamin D deficiency, respectively
abline(v = c(20,30), lty = 3)



## diagram, IRON in groups C0, C1 and C2
dfIron = data.frame(Iron = c(cases$iron, cases$iron2, controls$iron), 
                    group = c(rep("C1", length(cases$iron)), 
                              rep("C2", length(cases$iron2)), 
                              rep("C0", length(controls$iron)))
) 
boxplot(dfIron$Iron ~ dfIron$group, horizontal = T, main = "Iron", 
        xlab = "mcg/dl", ylab = "")
stripchart(dfIron$Iron ~ dfIron$group, pch = c(15, 16, 17), add = T,
           col = gray.colors(3, start = 0.1, end = 0.55, gamma = 3.2, 
                             rev = TRUE), cex = 1.5,
           method = "jitter", xlim=c(0,max(dfIron$Iron)+1), 
           main = "Iron values")

# line indicating iron deficiency
abline(v = c(50), lty = 3)


## diagram, FERRITIN in groups C0, C1 and C2
dfFerr = data.frame(Ferr = c(cases$ferritin, cases$ferritin2, controls$ferritin), 
                    group = c(rep("C1", length(cases$ferritin)), 
                              rep("C2", length(cases$ferritin2)), 
                              rep("C0", length(controls$ferritin)))
) 


boxplot(dfFerr$Ferr ~ dfFerr$group, horizontal = T, main = "Ferritin", 
        xlab = "ng/ml", ylab = "")
stripchart(dfFerr$Ferr ~ dfFerr$group, pch = c(15, 16, 17), add = T,
           col = gray.colors(3, start = 0.1, end = 0.55, gamma = 3.2, 
                             rev = TRUE), cex = 1.5,
           method = "jitter", xlim=c(0,max(dfIron$Iron)+1), 
           main = "Ferritin values")

# line indicating ferritin deficiency
abline(v = c(20), lty = 3)
par(mfrow = c(1,1))
```



There are unusually large values for `vitamin D` and `ferritin` in group C2. We next check whether these values belong to the same individual.


```{r}
which.max(cases$vitaminD2)
which.max(cases$ferritin2)
```

It is not the same individual. 


We check also whether these values are strong outliers. That is to say, whether they are beyond three times the IQR from the third quartile.
```{r}
# VITAMIN D, GROUP C2
max(cases$vitaminD2)
unname(quantile(cases$vitaminD2, probs = 0.75)+3*IQR(cases$vitaminD2))
```
Yes, it is a strong outlier. 

```{r}
# FERRITIN, GROUP C1
max(cases$ferritin)
quantile(cases$ferritin, probs = 0.75)+3*IQR(cases$ferritin)
```

No, it is not a strong outlier.

```{r}
# GROUP C2
max(cases$ferritin2)
quantile(cases$ferritin2, probs = 0.75)+3*IQR(cases$ferritin2)
```

No, it is not a strong outlier.


We decide do not remove the strong outlier for Vitamin D in group C2 even if it may affect the mean value due to the small sample size. When necessary, we will check results considering and not considering the strong outlier in order to assess its potential impact:

```{r}
cases_VitD = cases[ #cases$vitaminD2 != max(cases$vitaminD2)
                    , c("vitaminD", "vitaminD2", "weight", "age")]
```




### Dealing with potential confounders: age and weight
Note that `age` ranges from 2 to 13 years 

```{r}
range(c(cases$age, controls$age))
```

and `weight` from 10 to 69 kilograms

```{r}
range(c(cases$weight, controls$weight), na.rm = T)
```
To prevent confounding variables we analysed the possible association of `vitamin D`, `iron` and `ferritin` to the variables `age` or `weight` by testing H0: _"the slope of the regression line = 0 for each group and variable"_. Let us recall that if the linear assumptions are met (see below),
\[\Xi = \dfrac{b_1-\beta_1}{\sqrt{\dfrac{MEC}{(n-2)s^2(x)}}} \sim t_{n-2}\]
where $\beta_1$ is the slope of the linear model $y = \beta_0+\beta_1 x +\varepsilon$ and $b_1$ is its estimate, and $n$ is the sample size.

The following function calculates the value of $\Xi$ under the hypothesis $\beta_1=0$

```{r}
slopeRtest = function(x, y){ 
  # remove NAs
  not.na <- (!is.na(x))*(!is.na(y))
  x = x[which(not.na == 1)]
  y = y[which(not.na == 1)]
  
  # calculate regression line
  lmXY = lm(y ~ x)
  
  # calculate the statistic
  xi = unname(lmXY$coefficients[2])/( sum((lmXY$fitted.values - y)^2)/((length(x)-2)*var(x)))
  
  # calculate p-value
  min(2*pt(xi, df = length(x)-2), 
      2*pt(xi, df = length(x)-2, lower.tail = F)
      )
}
```

We next display the scatterplot along with the regression line, calculate the 95% confidence interval for $\beta_1$ (we use the `smatr` package ) and test $H_0:\,\it{\beta_1=0}$ (with the above defined function) for age, weight versus `vitamin D`, `iron` and `ferritin` for each group C0, C1 and C2.



In what follows,  `lm_XXX_y_z` refers to the linear model of variables `XXX`  over `y` in group `Cz`.


##### Vitamin D
Let us recall that we removed an strong outlier from groups C1 and C2 when dealing with `vitamin D`.



**Age**
```{r, fig.align='center', fig.width=9, fig.height=2}
lm_VitD_a_0 = lm(controls$vitaminD ~ controls$age)
lm_VitD_a_1 = lm(cases_VitD$vitaminD ~ cases_VitD$age)
lm_VitD_a_2 = lm(cases_VitD$vitaminD2 ~ cases_VitD$age)

par(mfrow = c(1,3))
plot(controls$age, controls$vitaminD, pch = 19, 
     xlab = "age", ylab = "Vitamin D", main = "C0")
abline(lm_VitD_a_0, col = "red", lwd = 1.5)

plot(cases_VitD$age, cases_VitD$vitaminD, pch = 19, 
     xlab = "age", ylab = "Vitamin D", main = "C1")
abline(lm_VitD_a_1, col = "red", lwd = 1.5)

plot(cases_VitD$age, cases_VitD$vitaminD2, pch = 19, 
     xlab = "age", ylab = "Vitamin D", main = "C0")
abline(lm_VitD_a_2, col = "red", lwd = 1.5)
par(mfrow = c(1,1))
```

```{r}
library(smatr)
confint(lm_VitD_a_0)[2, ]
slopeRtest(controls$age, controls$vitaminD)

confint(lm_VitD_a_1)[2, ]
slopeRtest(cases_VitD$age, cases_VitD$vitaminD)

confint(lm_VitD_a_2)[2, ]
slopeRtest(cases_VitD$age, cases_VitD$vitaminD2)
```

**Weight**
```{r, fig.align='center', fig.width=9, fig.height=2}
lm_VitD_w_0 = lm(controls$vitaminD ~ controls$weight)
lm_VitD_w_1 = lm(cases_VitD$vitaminD ~ cases_VitD$weight)
lm_VitD_w_2 = lm(cases_VitD$vitaminD2 ~ cases_VitD$weight)

par(mfrow = c(1,3))
plot(controls$weight, controls$vitaminD, pch = 19, 
     xlab = "weight", ylab = "Vitamin D", main = "C0")
abline(lm_VitD_w_0, col = "red", lwd = 1.5)

plot(cases_VitD$weight, cases_VitD$vitaminD, pch = 19, 
     xlab = "weight", ylab = "Vitamin D", main = "C1")
abline(lm_VitD_w_1, col = "red", lwd = 1.5)

plot(cases_VitD$weight, cases_VitD$vitaminD2, pch = 19, 
     xlab = "weight", ylab = "Vitamin D", main = "C0")
abline(lm_VitD_w_2, col = "red", lwd = 1.5)
par(mfrow = c(1,1))
```

```{r}
library(smatr)
confint(lm_VitD_w_0)[2, ]
slopeRtest(controls$weight, controls$vitaminD)

confint(lm_VitD_w_1)[2, ]
slopeRtest(cases_VitD$weight, cases_VitD$vitaminD)

confint(lm_VitD_w_2)[2, ]
slopeRtest(cases_VitD$weight, cases_VitD$vitaminD2)
```


##### `iron`

**`age``**
```{r, fig.align='center', fig.width=9, fig.height=2}
lm_iron_a_0 = lm(controls$iron ~ controls$age)
lm_iron_a_1 = lm(cases$iron ~ cases$age)
lm_iron_a_2 = lm(cases$iron2 ~ cases$age)

par(mfrow = c(1,3))
plot(controls$age, controls$iron, pch = 19, 
     xlab = "age", ylab = "Iron", main = "C0")
abline(lm_iron_a_0, col = "red", lwd = 1.5)

plot(cases$age, cases$iron, pch = 19, 
     xlab = "age", ylab = "Iron", main = "C1")
abline(lm_iron_a_1, col = "red", lwd = 1.5)

plot(cases$age, cases$iron2, pch = 19, 
     xlab = "age", ylab = "Iron", main = "C0")
abline(lm_VitD_a_2, col = "red", lwd = 1.5)
par(mfrow = c(1,1))
```

```{r}
library(smatr)
confint(lm_iron_a_0)[2, ]
slopeRtest(controls$age, controls$iron)

confint(lm_iron_a_1)[2, ]
slopeRtest(cases$age, cases$iron)

confint(lm_iron_a_2)[2, ]
slopeRtest(cases$age, cases$iron2)
```

**`weight`**
```{r, fig.align='center', fig.width=9, fig.height=2}
lm_iron_w_0 = lm(controls$iron ~ controls$weight)
lm_iron_w_1 = lm(cases$iron ~ cases$weight)
lm_iron_w_2 = lm(cases$iron2 ~ cases$weight)

par(mfrow = c(1,3))
plot(controls$weight, controls$iron, pch = 19, 
     xlab = "weight", ylab = "Iron", main = "C0")
abline(lm_iron_w_0, col = "red", lwd = 1.5)

plot(cases$weight, cases$iron, pch = 19, 
     xlab = "weight", ylab = "Iron", main = "C1")
abline(lm_iron_w_1, col = "red", lwd = 1.5)

plot(cases$weight, cases$iron2, pch = 19, 
     xlab = "weight", ylab = "Iron", main = "C0")
abline(lm_iron_w_2, col = "red", lwd = 1.5)
par(mfrow = c(1,1))
```

```{r}
library(smatr)
confint(lm_iron_w_0)[2, ]
slopeRtest(controls$weight, controls$iron)

confint(lm_iron_w_1)[2, ]
slopeRtest(cases$weight, cases$iron)

confint(lm_iron_w_2)[2, ]
slopeRtest(cases$weight, cases$iron2)
```




##### `ferritin`

**`age`**
```{r, fig.align='center', fig.width=9, fig.height=2}
lm_ferritin_a_0 = lm(controls$ferritin ~ controls$age)
lm_ferritin_a_1 = lm(cases$ferritin ~ cases$age)
lm_ferritin_a_2 = lm(cases$ferritin2 ~ cases$age)

par(mfrow = c(1,3))
plot(controls$age, controls$ferritin, pch = 19, 
     xlab = "age", ylab = "ferritin", main = "C0")
abline(lm_ferritin_a_0, col = "red", lwd = 1.5)

plot(cases$age, cases$ferritin, pch = 19, 
     xlab = "age", ylab = "ferritin", main = "C1")
abline(lm_ferritin_a_1, col = "red", lwd = 1.5)

plot(cases$age, cases$ferritin2, pch = 19, 
     xlab = "age", ylab = "ferritin", main = "C0")
abline(lm_VitD_a_2, col = "red", lwd = 1.5)
par(mfrow = c(1,1))
```

```{r}
library(smatr)
confint(lm_ferritin_a_0)[2, ]
slopeRtest(controls$age, controls$ferritin)

confint(lm_ferritin_a_1)[2, ]
slopeRtest(cases$age, cases$ferritin)

confint(lm_ferritin_a_2)[2, ]
slopeRtest(cases$age, cases$ferritin2)
```

**`weight`**
```{r, fig.align='center', fig.width=9, fig.height=2}
lm_ferritin_w_0 = lm(controls$ferritin ~ controls$weight)
lm_ferritin_w_1 = lm(cases$ferritin ~ cases$weight)
lm_ferritin_w_2 = lm(cases$ferritin2 ~ cases$weight)

par(mfrow = c(1,3))
plot(controls$weight, controls$ferritin, pch = 19, 
     xlab = "weight", ylab = "ferritin", main = "C0")
abline(lm_ferritin_w_0, col = "red", lwd = 1.5)

plot(cases$weight, cases$ferritin, pch = 19, 
     xlab = "weight", ylab = "ferritin", main = "C1")
abline(lm_ferritin_w_1, col = "red", lwd = 1.5)

plot(cases$weight, cases$ferritin2, pch = 19, 
     xlab = "weight", ylab = "ferritin", main = "C0")
abline(lm_ferritin_w_2, col = "red", lwd = 1.5)
par(mfrow = c(1,1))
```

```{r}
library(smatr)
confint(lm_ferritin_w_0)[2, ]
slopeRtest(controls$weight, controls$ferritin)

confint(lm_ferritin_w_1)[2, ]
slopeRtest(cases$weight, cases$ferritin)

confint(lm_ferritin_w_2)[2, ]
slopeRtest(cases$weight, cases$ferritin2)
```


#### Linear model assumptions

We use the `gvlma` package to assess residuals normality (skewness and kurtosis) and homocedasticity. The linear model assumptions are met in all of them but in few cases. When not met, either the p-value is close to 0.05, or only one of the normality assumptions is violated (recall that linear models are robust in front of loss of normality).

##### Vitamin D

```{r}
library(gvlma)
gvlma(lm_VitD_a_0)
gvlma(lm_VitD_a_1)
gvlma(lm_VitD_a_2)

gvlma(lm_VitD_w_0)
gvlma(lm_VitD_w_1)
gvlma(lm_VitD_w_2)
```

##### `iron`
```{r}
library(gvlma)
gvlma(lm_iron_a_0)
gvlma(lm_iron_a_1)
gvlma(lm_iron_a_2)

gvlma(lm_iron_w_0)
gvlma(lm_iron_w_1)
gvlma(lm_iron_w_2)
```

##### `ferritin`
```{r}
library(gvlma)
gvlma(lm_ferritin_a_0)
gvlma(lm_ferritin_a_1)
gvlma(lm_ferritin_a_2)

gvlma(lm_ferritin_w_0)
gvlma(lm_ferritin_w_1)
gvlma(lm_ferritin_w_2)
```

#### Normality of the variables
Cases (groups C1 and C2) are quite a little (19 or 20 individuals) so that normality must be checked to decide whether to apply a parametric test (`t.test`) or a non-parametric test (`wilcox_test` from the `coin` R package).

```{r}
# VITAMIN D
# C0
shapiro.test(controls$vitaminD)

# C1
shapiro.test(cases_VitD$vitaminD)

# C2
shapiro.test(cases_VitD$vitaminD2)
```

Normality is reached for groups C0 and C1, but not for group C2. 
Let us recall the existence of a strong outlier

```{r, fig.align='center', fig.height=3}
boxplot(cases_VitD$vitaminD2, horizontal = T)
stripchart(cases_VitD$vitaminD2, method = "jitter", cex = .5, pch = 19, add = T)
```

We test normality when removing the strong outlier with the Shapiro test 

```{r, fig.align='center', fig.height=3}
shapiro.test(cases_VitD$vitaminD2[cases_VitD$vitaminD2!= max(cases_VitD$vitaminD2)])
```

and also try with the Lilliefors (Kolmogorov-Smirnov) normality test qhen Shapiro's test is significant

```{r, fig.align='center', fig.height=3}
library(nortest)
lillie.test(cases_VitD$vitaminD2[cases_VitD$vitaminD2!= max(cases_VitD$vitaminD2)])
```

Both tests are either significant and yield p-values cole to the threshold value 0.05. 

We carry on by assuming that normality is not reached in this case. 



```{r, fig.align='center', fig.height=3}
##############################
# IRON
# C0
shapiro.test(controls$iron)

# C1
shapiro.test(cases$iron)

# C2
shapiro.test(cases$iron2)
```


```{r, fig.align='center', fig.height=3}
##############################
# FERRITIN
# C0
shapiro.test(controls$ferritin)
lillie.test(controls$ferritin)

# C1
shapiro.test(cases$ferritin)


# C2
shapiro.test(cases$ferritin2)
shapiro.test(cases$ferritin2[cases$ferritin2 != max(cases$ferritin2)])
lillie.test(cases$ferritin2)
```



## Calculations leading to Table 5 in the manuscript
### Compare `vitamin D`, `iron` and `ferritin` between groups C1 and C0


```{r}
#################################################
# VITAMIN D C1 vs C0
(vt = var.test(cases_VitD$vitaminD, controls$vitaminD, alternative = "two.sided"))
t.test(cases_VitD$vitaminD, controls$vitaminD, alternative = "less", var.equal = F)

mean(cases_VitD$vitaminD)
sd(cases_VitD$vitaminD)

mean(controls$vitaminD)
sd(controls$vitaminD)



# IRON C1 vs C0
var.test(cases$iron, controls$iron, alternative = "two.sided")
t.test(cases$iron, controls$iron, alternative = "less", var.equal = T)

mean(cases$iron)
sd(cases$iron)

mean(controls$iron)
sd(controls$iron)


# FERRITIN C1 vs C0
library(coin)
df10_ferr = data.frame(ferritin = c(cases$ferritin, controls$ferritin), 
                       group = as.factor(c(rep("C1", length(cases$ferritin)), 
                                 rep("C0", length(controls$ferritin))))
                       )
wilcox_test(ferritin ~ group, data = df10_ferr, distribution = "exact", conf.int = T, alternative = "less")
mean(cases$ferritin)
sd(cases$ferritin)

mean(controls$ferritin)
sd(controls$ferritin)
```



## Calculations leading to Table 6 in the manuscript
### Compare Vitamin D, `iron` and Ferrtin between groups C2 and C1
```{r}
# VITAMIN D C1 vs C2
var.test(cases_VitD$vitaminD2, cases_VitD$vitaminD, alternative = "two.sided")
t.test(cases_VitD$vitaminD2, cases_VitD$vitaminD, alternative = "greater", var.equal = F, paired = T)

library(coin)
df = stack(cases_VitD[, c("vitaminD2", "vitaminD")])[, c(2,1)]
wilcox_test(values ~ ind, data = df, distribution = "exact", conf.int = T)

df = stack(cases[, c("vitaminD2", "vitaminD")])[ , c(2,1)]
wilcox_test(values ~ ind, data = df, distribution = "exact", conf.int = T)

mean(cases_VitD$vitaminD)
sd(cases_VitD$vitaminD)

mean(cases_VitD$vitaminD2)
sd(cases_VitD$vitaminD2)


# IRON C2 vs C1
var.test(cases$iron2, cases$iron, alternative = "two.sided")
t.test(cases$iron2, cases$iron, paired = TRUE, alternative = "greater", var.equal = F)

mean(cases$iron)
sd(cases$iron)

mean(cases$iron2)
sd(cases$iron2)


# FERRITIN
# C2 vs C1
var.test(cases$ferritin2, cases$ferritin, alternative = "two.sided")
t.test(cases$ferritin2, cases$ferritin, paired = TRUE, alternative = "less", var.equal = T)

mean(cases$ferritin)
sd(cases$ferritin)

mean(cases$ferritin2)
sd(cases$ferritin2)
```


## Calculations leading to Table 7 in the manuscript
### Compare Vitamin D, `iron` and Ferrtin between groups C2 and C0
```{r}
# VITAMIN D C2 vs C0
var.test(cases_VitD$vitaminD2, controls$vitaminD, alternative = "two.sided")
t.test(cases_VitD$vitaminD2, controls$vitaminD, alternative = "two.sided", var.equal = F)


library(coin)

dfc2c0 = data.frame(values = c(cases_VitD$vitaminD2, controls$vitaminD),
                     ind = as.factor                  (c(rep("C2",length(cases_VitD$vitaminD2)),
                             rep("C0", length(controls$vitaminD)))))
wilcox_test(values ~ ind, data = dfc2c0, distribution = "exact", conf.int = T)

# IRON C2 vs C0
var.test(cases$iron2, controls$iron, alternative = "two.sided")
t.test(cases$iron2, controls$iron, alternative = "two.sided", var.equal = T)
mean(cases$iron2)
sd(cases$iron2)

mean(controls$iron)
sd(controls$iron)


# FERRITIN C2 vs C0
df02_ferr = data.frame(ferritin = c(cases$ferritin2, controls$ferritin), 
                       group = as.factor(c(rep("C2", length(cases$ferritin2)), rep("C0", length(controls$ferritin))))
                       )
wilcox_test(ferritin ~ group, data = df10_ferr, distribution = "exact", conf.int = T, alternative = "two.sided")

mean(cases$ferritin2)
sd(cases$ferritin2)

mean(controls$ferritin)
sd(controls$ferritin)
```



